\section{Introduction}
With the advent of new techniques of labeling and identifying cells, cortical wiring is no longer thought to be unorganized. Various rules governing the organization of synaptic connectivity have been proposed, such as dependency of connection probability and strength based on functional properties.  Specific connections between different cell types have been identified. Several types of ordered connectivity schemes have been discovered at the level of microcircuits, such as motifs formed due to strongly connected groups of neurons, groups of neurons specifically targeted by feed forward input. The local network architecture of pyramidal cells in the cortex are significantly different from an Erdös-Rényi network. The pyramidal neurons in the cortex have been found to be connected to each other in a non-random fashion forming structured local cortical circuits. The experimental data show that there are significant number of reciprocal connections between the pyramidal neurons and that there is a prevalence of other motifs involving groups of three or more highly interconnected neurons \cite{markram1997,thomson2002, Song2005, Perin2011}. Fast spiking interneurons and pyramidal cells are reciprocally connected with probability close to one \cite{Yoshimura2005}.\\
Together, these results provide a picture of local cortical microcircuit where specific connections occur greater than expected in a random network and a very few strong connections dominate in a sea of weak synapses. What then is the role of fine structure? How does it shape the dynamics and functional properties of cortical networks? One would expect that all the effects of adding structure to a random network at the microcircuit level would average out as the number of neurons are increased.  

%The local network architecture of pyramidal cells in the cortex are significantly different from an Erdös-Rényi network. The pyramidal neurons in the cortex have been found to be connected to each other in a non-random fashion forming structured local cortical circuits. The experimental data show that there are significant number of reciprocal connections between the pyramidal neurons and that there is a prevalence of other motifs involving groups of three or more highly interconnected neurons \cite{markram1997,thomson2002, Song2005, Perin2011}. Fast spiking interneurons and pyramidal cells are reciprocally connected with probability close to one \cite{Yoshimura2005}. 

Here we studied the simplest type of fine structure i.e. bidirectionality in the framework of Balanced networks \cite{carl1996, carl1998, carl2004}. This framework has been successful in explaining several ubiquitous features of cortical dynamics without resorting to fine tuning of network parameters \cite{softky1993, Holt1996, roxin2011}. These networks automatically find an operating point where total excitatory and inhibitory inputs to the neurons in the network approximately cancel or balance each other. The activity and consequently the functional properties of the network is then dictated by the modulations in the net input that are of the same order of magnitude as the threshold. In this study, we are interested in investigating if bidirectionality can lead to significant changes in the net input 


%Hence, the effective net inputs that the neurons receive are of same order of magnitude as the threshold $\left( O(Threshold) \right)$. Modulations or changes in the inputs that are $O(Threshold)$ lead to emergence of interesting network dynamics. In this study, we are interested in investigating if bidirectionality can lead to significant $O(Threshold)$ changes in the total input to neurons, and if so, to characterize the resulting changes on the network dynamics. In the standard balanced network the connectivity is assumed to be random, neurons receive weakly time-correlated order of threshold inputs or fast fluctuating inputs. As a result, input currents to the neurons in the network and consequently the spike time auto-correlations decorrelate rapidly\cite{carl1996, carl1998}. However, this is not consistent with what has been observed in cortical recordings, the power spectrum of membrane voltages recorded in the cortex show significant low frequency components. \cite{Tan2014}.	
	
We studied the effect of bidirectionality in a two population network consisting of Excitatory(E) and Inhibitory(I) neurons. We considered bidirectionality within each population(E-to-E and, I-to-I) and between the two populations(E-to-I) to address the following questions: What are the effects of excess bidirectionality on the on functional properties? What are its effects on the dynamics in a balanced network and on fluctuations and spiking statistics? We investigated these questions using numerical simulations of large scale spiking network  model of rodent V1 (see Methods). \\
